Empowering the use of community driven preprints in ecology and evolution

### 2024-04-14 ms/ms.qmd
Fix fig 1

We were interested in getting an understanding of whether career stage is related to preprint use. We used the first publication year of the submitting author as a proxy for career stage. We predicted that early career researchers would be more open to using preprints in Ecology and Evolution. Interestingly, there was strong evidence that the number of preprints posted decreases with later career stages (negative binomial glm: year slope = `r res[2, "Estimate"]`, SE: `r res[2, "Std. Error"]`, *p* < 0.001, N = `r nrow(count_dat[complete.cases(count_dat),])`). The majority of preprints were submitted by authors who published their first paper in the last 10 years (@fig-career), with the median year of first publication of submitting authors being `r median(data2$submitting_author_first_publication_year, na.rm = TRUE)` (mean = `r round(mean(data2$submitting_author_first_publication_year, na.rm = TRUE),0)`; SD = `r sd(data2$submitting_author_first_publication_year, na.rm = TRUE)`).

Gender differences in preprint use have been established in a number of other research fields. It is hypothsised that such differences are driven by the larger childcare burden of female researchers, which limit their abilities to publish as frequently as their male counterparts [@ucar2022mind; @wehner2020comparison]. Recent work has shown that COVID-19 exacerbated gender disparities in preprint publishing given the increased childcare burden during the pandemic associated with lockdowns and home schooling [@wehner2020comparison].

Understanding gender publishing patterns is challenging. We cannot know the gender of authors for certainty but we can use a data driven approach to suggest the probability that a particular name is of a given gender (male or female). We used the R package *DemografixeR* (vers. `r utils::packageVersion("DemografixeR")`; @RJ-2016-002) to predict the most likely gender of the submitting author of a preprint. We incorperated information about the submitting authors country of affiliation to better account for differences across countries in naming.

@@

Despite the early successes of new initiatives taken by *EcoEvoRxiv* it is still clear that much work needs to be done to improve authors' understanding of preprints (and postprints). Our community needs to work harder to help educate our colleagues in this space. We view this perspective piece as one small step in achieving that goal. Our data driven approach has allowed us to better understand some of the key elements of uncertainty surrounding preprints which provides us with opportunties to clarify our perspective and recommendations. First, authors need not be afraid of sharing data and code early in the publication process. Sharing data and code early can help improve the quality of research, establish precedence, and improve the transparency and computational reproducibility of scientific findings. If authors are worried about data being used unintentially, clear information surrounding its re-use and withholding some data or meta-data needed to understand the dataset may be options. Second, authors need to be aware of the benefits of posting preprints and postprints. Posting preprints can help establish precedence of research findings, ensure that research is open access, and help authors receive feedback on their research before it is published. Third, authors need to be aware of the benefits of using community driven preprint servers. Community driven preprint servers can help ensure that research is open access, that research is reviewed by experts in the field, and that research is disseminated quickly.

>>>>

We were interested in getting an understanding of whether career stage is related to preprint use. We used the first publication year of the submitting author as a proxy for career stage. We predicted that early career researchers would be more open to using preprints in Ecology and Evolution. Interestingly, there was strong evidence that the number of preprints posted decreases with later career stages (negative binomial glm: year slope = `r res[2, "Estimate"]`, SE: `r res[2, "Std. Error"]`, *p* < 0.001, N = `r nrow(count_dat[complete.cases(count_dat),])`). The majority of preprints were submitted by authors who published their first paper in the last 10 years ([@fig-summary]E), with the median year of first publication of submitting authors being `r median(data2$submitting_author_first_publication_year, na.rm = TRUE)` (mean = `r round(mean(data2$submitting_author_first_publication_year, na.rm = TRUE),0)`; SD = `r sd(data2$submitting_author_first_publication_year, na.rm = TRUE)`).

Gender differences in preprint use have been established in a number of other research fields. It is hypothsised that such differences are driven by the larger childcare burden of female researchers, which limit their abilities to publish as frequently as their male counterparts [@ucar2022mind; @wehner2020comparison]. Understanding gender publishing patterns is challenging. We cannot know the gender of authors for certainty but we can use a data driven approach to suggest the probability that a particular name is of a given gender (male or female). We used the R package *DemografixeR* (vers. `r utils::packageVersion("DemografixeR")`; @RJ-2016-002) to predict the most likely gender of the submitting author of a preprint. We incorperated information about the submitting authors country of affiliation to better account for differences across countries in naming.

@@

Despite the early successes of new initiatives taken by *EcoEvoRxiv* it is still clear that much work needs to be done to improve our communities understanding of preprints (and postprints). Our community needs to work harder to help educate colleagues in this space. We view this perspective piece as one small step in achieving that goal.

Our data driven approach has allowed us to better understand some of the key elements of uncertainty surrounding preprints which provides us with opportunties to clarify our perspective and recommendations. First, authors need not be afraid of sharing data and code early in the publication process. Sharing data and code early can help improve the quality of research, establish precedence, and improve the transparency and computational reproducibility of scientific findings. If authors are worried about data being used unintentially, clear information surrounding its re-use and withholding some data or meta-data needed to understand the dataset may be options. Second, authors need to be aware of the benefits of posting preprints and postprints. Posting preprints can help establish precedence of research findings, ensure that research is open access, and help authors receive feedback on their research before it is published. Third, authors need to be aware of the benefits of using community driven preprint servers. Community driven preprint servers can help ensure that research is open access, that research is reviewed by experts in the field, and that research is disseminated quickly.
